---
title: "Visualizing Starwars characters"
author: "Mine Çetinkaya-Rundel"
format: html
editor: visual
---

```{r}
#| label: load-packages
#| include: false

library(tidyverse)
```

1.  Glimpse at the starwars data frame.

```{r}
#| label: glimpse-starwars

glimpse(starwars)
```

2.  Modify the following plot to change the color of all points to `"pink"`.

```{r}
#| label: scatterplot

ggplot(starwars, 
       aes(x = height, y = mass, color = gender, size = birth_year)) +
  geom_point(color = "pink")
```

3.  Add labels for title, x and y axes, and size of points. Uncomment to see the effect.

```{r}
#| label: scatterplot-labels

ggplot(starwars, 
       aes(x = height, y = mass, color = gender, size = birth_year)) +
  geom_point(color = "#30509C") +
  labs(
    title = "Height vs Mass of Starwars Characters",
    x = "Height (cm)", 
    y = "Mass (kg)",
    size = "Birth Year"
    )
```

\

4.  Pick a single numerical variable and make a histogram of it. Select a reasonable binwidth for it.

(A little bit of starter code is provided below, and the code chunk is set to not be evaluated with `eval: false` because the current code in there is not valid code and hence the document wouldn't knit. Once you replace the code with valid code, set the chunk option to `eval: true`, or remove the `eval` option altogether since it's set to `true` by default.)

```{r}
#| label: histogram
#| eval: false

ggplot(starwars, aes(height)) +
  geom_histogram(bins = 40)
```

5.  Pick a numerical variable and a categorical variable and make a visualization (you pick the type!) to visualization the relationship between the two variables. Along with your code and output, provide an interpretation of the visualization.

```{r}
#| label: num-cat

# A boxplot can have a categorical and numerical variable on either axis
ggplot(starwars, aes(x = sex, y = height)) +
  geom_boxplot()

ggplot(starwars, aes(x = gender, y = height)) +
  geom_boxplot()

# A density plot can take a numerical and categorical value to colour by the categorical variable.
ggplot(starwars, aes(x = height, colour = sex)) +
  geom_density()
```

```{r}
#| label: num-cat
starwarsg <- starwars |>
  drop_na()
```

Interpretation goes here...

6.  Pick a single categorical variable from the data set and make a bar plot of its distribution.

```{r}
#| label: barplot

ggplot(starwars, aes(x=gender)) +
  geom_bar(fill="red")

ggplot(starwars, aes(x=gender)) +
  geom_bar(color="red")

```

```{r}

#| label: barplot

ggplot(starwars, aes(x=species)) +
  geom_bar(fill="red")

ggplot(starwars, aes(x=sex)) +
  geom_bar(color="red")

```

Pick two categorical variables and make a visualization to visualize the relationship between the two variables. Along with your code and output, provide an interpretation of the visualization.

```{r}
#| label: cat-cat
ggplot(starwars, aes(x=gender, fill=sex)) +
  geom_bar()
```

Interpretation goes here...

8.  Pick two numerical variables and two categorical variables and make a visualization that incorporates all of them and provide an interpretation with your answer.

(This time no starter code is provided, you're on your own!)

```{r}
#| label: multi
ggplot(starwars, aes(x = height, y = birth_year)) +
  geom_point(aes(color = sex, shape = gender))
```

Interpretation goes here...
