---
title: "Lab 01 - Hello R"
author: "Kedman"
date: "2/12/2025"
output: github_document
---

## Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
library(datasauRus)
```

## Exercises

### Exercise 1

```{r}
datasaurus_dozen %>% 
  count(dataset) 
```

### Exercise 2


First let's plot the data in the dino dataset:

```{r plot-dino, fig.height=3, fig.width=6}
dino_data <- datasaurus_dozen %>%
  filter(dataset == "dino")

ggplot(data = dino_data, mapping = aes(x = x, y = y)) +
  geom_point()
```

And next calculate the correlation between `x` and `y` in this dataset:

```{r cor-dino}
dino_data %>%
  summarize(r = cor(x, y))
```

### Exercise 3

```{r plot-star, fig.height=3, fig.width=6}
star_data <- datasaurus_dozen %>%   
filter(dataset == "star") 

ggplot(data = star_data, mapping = aes(x = x, y = y)) +
  geom_point()
```

```{r cor-star}
star_data %>% 
  summarise(r = cor(x, y))

```
The correlation coefficient for star-data (-0.06296) and dino-data-(-0.06447185) are similar as we see.

### Exercise 4

```{r plot-circle, fig.height=3, fig.width=6}
circle_data <- datasaurus_dozen %>%   
  filter(dataset == "circle")

ggplot(data = circle_data, mapping = aes(x = x, y = y)) +
  geom_point()

```

```{r cor-circle}
circle_data %>% 
  summarize(r = cor(x, y))

```
The correlation coefficient for circle-data (-0.06834336) and dino-data-(-0.06447185) are not so far apart.

### Exercise 5

Add code and narrative as needed.
To add R chunks either type out the backticks, curly braces, and the letter `r` or use the Insert chunk button above, green C+.
